# Regression Approach to Treatment Effect Estimation

Suppose one would like to use a regression model to estimate the treatment effect of a SAT, but controlling for the covariate 'SES'. 

Simulating data: 

```{r}
ID = c(1:6)
Grp = rep(c(0, 1), each = 3)
Score = c(550, 600, 650, 600, 720, 630)
SES = c(1, 2, 2, 2, 3, 2)

SATdat = data.frame(ID, Grp, Score, SES)
SATdat$SES = factor(SATdat$SES)
```

## Regression w/ no confounder

```{r}
lm1 = lm(Score ~ Grp, data = SATdat)
summary(lm1)
```

This is the same as a simple t-test: 

```{r}
t.test(SATdat$Score[SATdat$Grp==1], SATdat$Score[SATdat$Grp==0])
```

not sig -- sample size is too small

## Including confounder SES

```{r}
ses.lm = lm(Score ~ Grp+SES, data = SATdat)
summary(ses.lm)
```

The effect (-10) is assumed between groups WITHIN EACH SES LEVEL: 

### low SES 
```{r}
predict(ses.lm, newdata = data.frame(Grp=0, SES=factor(1)))
predict(ses.lm, newdata = data.frame(Grp=1, SES=factor(1)))
```

### middle SES 
```{r}
predict(ses.lm, newdata = data.frame(Grp=0, SES=factor(2)))
predict(ses.lm, newdata = data.frame(Grp=1, SES=factor(2)))
```

### high SES 
```{r}
predict(ses.lm, newdata = data.frame(Grp=0, SES=factor(3)))
predict(ses.lm, newdata = data.frame(Grp=1, SES=factor(3)))
```

