# Types of Causal Effects

Sometimes our experiments because quasi-experiments because of crossover or non-compliance. 

Simulating data on assignment, compliance, and potential outcomes: 

```{r}
set.seed(1)
assign = rbinom(100, 1, 0.5) # half half
```

Simulating non-compliance as a function of gender: 

```{r}
set.seed(101)
gender = rbinom(100, 1, 0.4) # 40% are men
comply = rep(); 
comply[gender==1] = rbinom(sum(gender==1), 1, 0.5) # prob of men complying is 0.5
comply[gender==0] = rbinom(sum(gender==0), 1, 0.7) # prob of women complying is 0.7
```

treatment effect: 100 (from 400 --> 500) fixed for everyone:
 
```{r}
notutor = rnorm(100, 400, 12) # Y0
tutor = notutor + 100 + rnorm(100, 0, 20) # Y1
simdata = data.frame(assign, notutor, tutor)
head(simdata)
```

actual treatment:

```{r}
treat = ifelse(comply==1, assign, 1-assign)
table(assign, treat)
```

observed value:

```{r}
obs = treat*tutor + (1-treat)*notutor # Y1 if T=1 and Y0 if T=0
```
* when treat=1: Y = 1*tutor + 0*notutor = Y1
* when treat=0: Y = 0*tutor + 1*notutor = Y0

Simulated data: 

```{r}
simdata = data.frame(assign, comply, treat, tutor, notutor, obs, gender)
head(simdata)
```


## ATT, ATE, ATU, ... what the what? 

* ATE/ACE: Average Treatment Effect in the population
* ATT: Average Treatment Effect on the Treated
* ATU: Average Treatment Effect on the Untreated
* LATE: Local Average Treatment Effect for Compilers
* CATE: Conditional Average Treatment Effect
* ITT: Intention To Treat. Effect of simply the intention (instead of receipt) of treatment

```{r}
ATE = mean(simdata$tutor - simdata$notutor)  # whole population
ATT = mean((simdata$tutor - simdata$notutor)[treat==1])
ATU = mean((simdata$tutor - simdata$notutor)[treat==0])
LATE = mean((simdata$tutor - simdata$notutor)[comply==1])
CATE0 = mean((simdata$tutor - simdata$notutor)[gender==0])
CATE1 = mean((simdata$tutor - simdata$notutor)[gender==1])
```

```{r}
cbind(ATE, ATT, ATU, LATE, CATE0, CATE1)
# all close to ATE
```

## Calculating different causal effects in practice

* ATE
```{r}
ATE_obs = lm(obs ~ treat, data = simdata)
coef(ATE_obs) # 99.82048 
```

* ATT & ATU

no way to estimate ATT & ATU from data

* LATE if compliance status is known

```{r}
LATE_obs = lm(obs ~ treat, data = simdata[simdata$comply==1,])
coef(LATE_obs) # 101.1594 
```

* CATE if gender is included as a confounder:

```{r}
CATE0_obs = lm(obs ~ treat, data = simdata[simdata$gender==0,])
coef(CATE0_obs) # 106.5414 

CATE1_obs = lm(obs ~ treat, data = simdata[simdata$gender==1,])
coef(CATE1_obs) # 90.71861 
```

* Intention To Treat (ITT)

```{r}
ITT = lm(obs ~ assign)
coef(ITT) # 15.460 whereas the true effect is 100!!
```

* Intention To Treat on Compliers

```{r}
ITT_c = lm(obs ~ assign, data = simdata[simdata$comply==1,])
coef(ITT_c) # 101.159 which is pretty close to the true effect 100
```


**Generally recommended: LATE or ITT on Compliers!** 




