# PSM Matching Strategy

```{r, message=FALSE}
library(Matching)
library(MatchIt)
library(optmatch)
library(weights)
library(cem)
library(tcltk2)
```

## Data

```{r data}
data(lalonde)
dim(lalonde)
names(lalonde)
```

### Create dummy variable for race

```{r}
lalonde$black = ifelse(lalonde$race=='black', 1, 0)
lalonde$hispan = ifelse(lalonde$race=='hispan', 1, 0)
```

## PSM Steps

### Selection of covariates in X

```{r}
fm1 = treat ~ age + educ + black + hispan + married + I(re74/1000) + I(re75/1000)
fm2 = treat ~ age + I(age^2) + I(age^3) + educ + black + hispan + married + I(re74/1000) + I(re75/1000)
fm3 = treat ~ age + I(age^2) + I(age^3) + educ + I(educ^2) + black + hispan + married + I(re74/1000) + I(re75/1000)
```

### Calculation of propensity scores (p-scores)
```{r}
pscore <- glm(fm1, data = lalonde, family = 'binomial')
head(pscore$fitted.values)
hist(pscore$fitted.values[lalonde$treat==0],xlim=c(0,1))
hist(pscore$fitted.values[lalonde$treat==1],xlim=c(0,1))
lalonde$pscore = pscore$fitted.values
```
  
try other formulas?

### Matching based on p-sores

all in one function matchit()

```{r}
set.seed(42)
m.out <- matchit(data = lalonde,
                 formula = fm1,
                 distance = "logit",
                 method = "nearest",
                 replace = TRUE,
                 caliper = 0.2, 
                 discard = 'both'
                 )
```

* m.out$match.matrix
* m.out$distance
* plot(m.out$distance, m.out$fitted.values) # same 
* method: exact, subclass, optimal, full, cem
* distance: pscore
* plot(m.out, type = "qq", interactive=FALSE)

```{r}
plot(summary(m.out))
```

Balance for formula 2: 

```{r}
set.seed(42)
m.out <- matchit(data = lalonde,
                 formula = fm2,
                 distance = "logit",
                 method = "nearest",
                 replace = TRUE,
                 caliper = 0.2, 
                 discard = 'both'
                 )
plot(summary(m.out))
```

Balance for formula 3: 

```{r}
set.seed(42)
m.out <- matchit(data = lalonde,
                 formula = fm3,
                 distance = "logit",
                 method = "nearest",
                 replace = TRUE,
                 caliper = 0.2, 
                 discard = 'both'
                 )
plot(summary(m.out))
```

* matchit: https://kosukeimai.github.io/MatchIt/articles/MatchIt.html
* manual: https://imai.fas.harvard.edu/research/files/matchit.pdf
* starting from page 15
* Let's play with the arguments! 

