# PSM Matching Strategy

```{r, message=FALSE}
library(Matching)
library(MatchIt)
library(optmatch)
library(weights)
library(cem)
library(tcltk2)
library(knitr)
library(CBPS)
library(jtools)
library(cobalt)
library(lmtest)
library(sandwich) #vcovCL
library(rbounds) #gamma
```

## Data

```{r}
#data(lalonde)
lalonde = MatchIt::lalonde
dim(lalonde)
names(lalonde)
```

### Create dummy variable for race and unemployment

```{r}
lalonde$black = ifelse(lalonde$race=='black', 1, 0)
lalonde$hispan = ifelse(lalonde$race=='hispan', 1, 0)
lalonde$un74 = ifelse(lalonde$re74==0, 1, 0)
lalonde$un75 = ifelse(lalonde$re75==0, 1, 0)
```

## Regression Estimates
### naive t-test estimate 

```{r}
reglm <- lm(re78 ~ treat, data = lalonde)
summ(reglm)
```

### regression with covariates 
```{r}
reglm1 <- lm(re78 ~ treat + educ + age + black + hispan + married + nodegree + un74 + un75 + re74 + re75, data = lalonde)
summ(reglm1)
```


## PSM Steps

### Selection of covariates in X

```{r}
fm1 = treat ~ age + educ + black + hispan + married + I(re74/1000) + I(re75/1000)
fm2 = treat ~ age + I(age^2) + I(age^3) + educ + black + hispan + married + I(re74/1000) + I(re75/1000)
fm3 = treat ~ age + I(age^2) + I(age^3) + educ + I(educ^2) + black + hispan + married + I(re74/1000) + I(re75/1000)
```

### Calculation of propensity scores (p-scores)
```{r}
pscore <- glm(fm2, data = lalonde, family = 'binomial')
head(pscore$fitted.values)
hist(pscore$fitted.values[lalonde$treat==0],xlim=c(0,1))
hist(pscore$fitted.values[lalonde$treat==1],xlim=c(0,1))
lalonde$pscore = pscore$fitted.values
```

Logistic regression:

```{r}
summ(pscore)
```

try other formulas?



### Matching based on p-sores

all in one function matchit():

```{r}
set.seed(42)
m.out <- matchit(data = lalonde,
                 formula = fm1,
                 distance = "logit",
                 method = "nearest",
                 replace = TRUE,
                 caliper = 0.2, 
                 discard = 'both'
                 )
```

summary of matching results:

```{r}
summary(m.out)
```

* m.out$match.matrix
* m.out$distance
* plot(m.out$distance, m.out$fitted.values) # same 
* method: exact, subclass, optimal, full, cem
* distance: pscore
* plot(m.out, type = "qq", interactive=FALSE)

### Checking balance on covariates:

#### Balance for formula 1: 

```{r}
plot(summary(m.out))
```


#### Balance for formula 2: 

```{r}
set.seed(42)
m.out <- matchit(data = lalonde,
                 formula = fm2,
                 distance = "logit",
                 method = "nearest",
                 replace = TRUE,
                 caliper = 0.2, 
                 discard = 'both'
                 )
plot(summary(m.out))
love.plot(m.out, binary = "std")
```

Standardized mean differences (SMD): 

```{r, eval=T}
summary(m.out)
#out = summary(m.out)
#round(out$sum.all, 3)
#round(out$sum.matched, 3)
```

```{r, eval=T}
plot(m.out, type = "hist", interactive = F)
plot(m.out, type = "jitter", interactive = F)
love.plot(m.out, binary = "std")
bal.plot(m.out, var.name = "distance", which = "both",
         type = "histogram", mirror = TRUE)
```

who matched to whom?

```{r}
head(m.out$match.matrix, 10)
```


#### Balance for formula 3: 

```{r}
set.seed(42)
m.out <- matchit(data = lalonde,
                 formula = fm3,
                 distance = "logit",
                 method = "nearest",
                 replace = FALSE,
                 caliper = 0.2, 
                 discard = 'both'
                 )
```

```{r}
summary(m.out)
```


```{r}
plot(summary(m.out))
love.plot(m.out, binary = "std")
```

### Estimation of treatment effect

I will use the matching results with fm3 because it yielded the best balance on the covariates between the matched groups. I also kept ratio = 1 and replace = FALSE to avoid the specification of weights for now: 

```{r}
set.seed(42)
m.out <- matchit(data = lalonde,
                 formula = fm3,
                 distance = "logit",
                 method = "nearest",
                 replace = TRUE,
                 ratio = 1, 
                 caliper = 0.2, 
                 discard = 'both',
                 estimand = 'ATC'
                 )
plot(summary(m.out))
love.plot(m.out, binary = "std")
```

Extract matched data: 

```{r}
m.data <- match.data(m.out)
# m.data <- get_matches(m.out)
tail(m.data)
dim(m.data)
```

#### Linear model without covariates:

```{r}
fit1 <- lm(re78 ~ treat, data = m.data)
summ(fit1)
```

Cluster-robust standard errors:

```{r}
coeftest(fit1, vcov. = vcovCL)
```

Alternatively, use weighted Student's t-test: 

```{r}
res <- wtd.t.test(m.data$re78[m.data$treat == 1],
                  m.data$re78[m.data$treat == 0],
                  weight = m.data$weights[m.data$treat == 1],
                  weighty = m.data$weights[m.data$treat == 0])
print(res)
mu <- res$additional[1]
std <- res$additional[4]
cat("Confidence interval: ", sapply(qt(c(0.025, 0.975), coef(res)["df"]), function(x){return(mu+x*std)}), "\n")
```

#### Linear model with covariates: double robust (use with caution)

```{r}
fit2 <- lm(re78 ~ treat + age + educ + black + hispan + married + nodegree + un74 + un75 + re74 + re75, data = m.data)
# include only the covariate that is not balanced
fit2 <- lm(re78 ~ treat + I(educ^2), data = m.data)
summ(fit2)
```

```{r}
coeftest(fit2, vcov. = vcovCL)
```

### Hidden bias analysis

extract matched treatment and control units from m.data

```{r}
psens(x = m.data$re78[m.data$treat==1], 
      y = m.data$re78[m.data$treat==0], 
      Gamma = 3, GammaInc=0.1)
```

a gamma value with 1.6 or larger could lead to a change in the ATT estimate.




### References:

* matchit: https://kosukeimai.github.io/MatchIt/articles/MatchIt.html
* manual: https://imai.fas.harvard.edu/research/files/matchit.pdf
* https://cran.r-project.org/web/packages/MatchIt/vignettes/assessing-balance.html 
* starting from page 15
* Let's play with the arguments! 

