# Propensity Score Analysis

In this lab, we will use propensity scores to perform other types of analyses (weighting, stratification, and covariate adjustment)

```{r, message=FALSE}
library(weights)
library(survey)
library(twang)
library(CBPS)
library(cobalt)
library(jtools)
library(lmtest)
library(sandwich) #vcovCL
library(rbounds) #gamma
library(tidyr)
library(tidyverse)
library(janitor)
#remotes::install_github("vdorie/treatSens")
#library(treatSens)
```

<!------------------------------>
## Data
<!------------------------------>

```{r}
#data(lalonde)
lalonde = MatchIt::lalonde
dim(lalonde)
names(lalonde)
```

### Create dummy variable for race and unemployment

```{r}
lalonde$black = ifelse(lalonde$race=='black', 1, 0)
lalonde$hispan = ifelse(lalonde$race=='hispan', 1, 0)
lalonde$un74 = ifelse(lalonde$re74==0, 1, 0)
lalonde$un75 = ifelse(lalonde$re75==0, 1, 0)
```

<!------------------------------>
## Calculating Propensity Scores
<!------------------------------>

### Different formulas

```{r}
fm1 = treat ~ age + educ + black + hispan + married + I(re74/1000) + I(re75/1000)
fm2 = treat ~ age + I(age^2) + I(age^3) + educ + black + hispan + married + I(re74/1000) + I(re75/1000)
fm3 = treat ~ age + I(age^2) + I(age^3) + educ + I(educ^2) + black + hispan + married + I(re74/1000) + I(re75/1000)
```

### Calculation of propensity scores (p-scores)
```{r}
pscore <- glm(fm3, data = lalonde, family = 'binomial')
head(pscore$fitted.values)
hist(pscore$fitted.values[lalonde$treat==0],xlim=c(0,1))
hist(pscore$fitted.values[lalonde$treat==1],xlim=c(0,1), add=F)
lalonde$pscore = pscore$fitted.values
```

```{r}
hist(pscore$fitted.values[lalonde$treat==0],xlim=c(0,1), density = 20, angle = 45, main="Propensity Scores", xlab="Shaded = Untreated | Gray = Treated")
hist(pscore$fitted.values[lalonde$treat==1],xlim=c(0,1), col=gray(0.4,0.25),add=T)
```

<!------------------------------>
## IPTW: Inverse Probability Treatment Weighting
<!------------------------------>

### weightATT

weightATT is created by using the ifelse() function to obtain:
* 1 for treated units
* p/(1-p) for control units

```{r}
lalonde$weightATT <- with(lalonde, ifelse(treat==1, 1, pscore/(1-pscore)))
```

A summary of the ATT weights for treated and untreated groups:

```{r}
with(lalonde, by(pscore,treat,summary)) 
```

```{r}
with(lalonde, by(weightATT,treat,summary)) 
boxplot(lalonde$weightATT ~ lalonde$treat)
```

### weightATE

weightATE is created by using the ifelse() function to obtain:
* 1/p for treated units
* 1/(1-p) for control units

```{r}
lalonde$weightATE <- with(lalonde, ifelse(treat==1, 1/pscore, 1/(1-pscore)))
```

A summary of the ATE weights for treated and untreated groups:

```{r}
with(lalonde, by(weightATE,treat,summary)) 
boxplot(lalonde$weightATE ~ lalonde$treat)
```

* the maximum weight for the ATE is 33.961 

### Weight truncation

* Truncation can be performed by assigning the weight at a cutoff percentile to observations with weights above the cutoff.

* The following code demonstrates weight truncation. More specifically, it uses the quantile function to calculate the weight at the 99th percentile and the ifelse function to assign this weight to any student whose weight exceeds the 99th percentile:

```{r}
lalonde$weightATETruncated <- with(lalonde, ifelse(weightATE > quantile(weightATE, 0.99), quantile(weightATE, 0.99), weightATE))
```

```{r}
with(lalonde, by(weightATETruncated,treat,summary)) 
boxplot(lalonde$weightATETruncated ~ lalonde$treat)
```

may truncate all above .05 quantiles if necessary

```{r}
with(lalonde, by(weightATETruncated,treat,summary)) 
boxplot(lalonde$weightATETruncated ~ lalonde$treat)
```


### Balance check

In this section, covariate balance evaluation is performed by comparing the standardized difference between the weighted means of the treated and untreated groups.

The bal.stat function of the twang package (Ridgeway et al., 2013) is useful for covariate balance evaluation. 

If there are no sampling weights, then sampw=1.

```{r}
covariateNames <-names(lalonde)[2:14]
balanceTable <- bal.stat(lalonde, vars= covariateNames,
                         treat.var = "treat",
                         w.all = lalonde$weightATE, 
                         get.ks=F,
                         sampw = 1,
                         estimand="ATE", multinom=F)
balanceTable <- balanceTable$results
round(balanceTable,3) 
```

* std.eff.sz quantifies effect size
* 0.2 small 0.5 medium 0.8 large
* most of them are small (except nodegree and un74) - balance was somewhat achieved *

### Estimation of treatment effect

the final weights are divided by the mean of weights to make them sum to the sample size, which is a process known as normalization.

```{r}
lalonde$finalWeight <- lalonde$weightATE/mean(lalonde$weightATE)
```

Before the estimation can be performed, the surveyDesign object is created with the svydesign function to declare the names of the variables that contain cluster ids, strata ids, weights, and the data set: 

```{r}
surveyDesign <- svydesign(ids=~1, weights=~finalWeight, data = lalonde, nest=T) 
```

Methods to obtain standard errors for propensity score weighted estimates include Taylor series linearization and resampling methods such as bootstrapping, jackknife, and balanced repeated replication (see review by Rodgers, 1999).

To use bootstrap methods with the survey package, the surveyDesign object created above should be modified to include weights for each replication. The following code takes the surveyDesign object and adds weights for 1,000 bootstrapped samples:

```{r}
set.seed(8)
surveyDesignBoot <- as.svrepdesign(surveyDesign, type=c("bootstrap"), replicates=1000)
```

First, the svyby function is used to apply the svymean function separately to treated and control units to obtain weighted outcome:

```{r}
weightedMeans <- svyby(formula=~re78, by=~treat, design=surveyDesignBoot, FUN=svymean, covmat=TRUE)
weightedMeans
```


*The code to obtain the treatment effect with a regression model is shown as follows. 
* The model formula using R notation is re78~treat. This formula can be expanded to include any covariates and interaction effects of interest

```{r}
outcomeModel <- svyglm(re78~treat, design = surveyDesignBoot)
summary(outcomeModel)
```
Double robust with the imbalanced covariates:

```{r}
outcomeModel <- svyglm(re78 ~ treat + age + un74 + hispan, design = surveyDesignBoot)
summary(outcomeModel)
```


<!------------------------------>
## Stratification
<!------------------------------>

### Manually creating subclasses

* The following code used the cut function to create five strata of approximately the same size based on the quintiles of the distribution of propensity scores for both treated and untreated groups. 
* The quintiles are obtained with the quantile function. The function levels is used to assign number labels from 1 to 5 to the strata, and then xtabs is used to display strata by treatment counts.
* The number of strata is limited by the common support of the propensity score distributions of treated and untreated groups, because each stratum must have at least one treated and one untreated observation.


```{r}
hist(lalonde$pscore)
quantile(lalonde$pscore, prob = seq(0, 1, 1/5))

lalonde$subclass <- cut(x=lalonde$pscore, breaks = quantile(lalonde$pscore, prob = seq(0, 1, 1/5)), include.lowest=T)
levels(lalonde$subclass) <- 1:length(levels(lalonde$subclass))

ntable <- xtabs(~treat+subclass,lalonde) 
ntable

surveyDesign <- svydesign(ids=~1, weights=~lalonde$finalWeight, 
                          data = lalonde, nest=T) 
```

### Stratification using matchit():

```{r}
set.seed(42)
stratification <- matchit(data = lalonde,
                 formula = fm2,
                 distance = "logit",
                 method = "subclass",
                 subclass = 5)
```


### Covariate Balance Evaluation

```{r}
balance.stratification = summary(stratification)
balance.stratification$qn
```
* It is noticeable in the cross-classification of treatment by strata shown above that the stratification based on the propensity scores of the treated resulted in a similar number of treated units within strata.

* When using summary(), the default is to display balance only in aggregate using the subclassification weights. This balance output looks similar to that for other matching methods.

```{r}
round(balance.stratification$sum.across,3)
```

* If the goal is to estimate the treatment effect by pooling stratum-specific treatment effects, covariate balance should be evaluated and achieved within strata. * However, if the number of covariates is large, evaluation of covariate balance within strata can become cumbersome. 
* Also, if the sample sizes of treated or untreated groups within strata are small, covariate balance evaluation can become very sensitive to outliers.

* An additional option in summary(), subclass, allows us to request balance for individual subclasses. 
* Below we call summary() and request balance to be displayed on all subclasses (setting un = FALSE to suppress balance in the original sample):

```{r}
summary(stratification, subclass = TRUE, un = FALSE)
```

* We can plot the standardized mean differences in a Love plot that also displays balance for the subclasses using plot.summary.matchit() on a summary.matchit() object with subclass = TRUE.


```{r}
balance.stratification2 <- summary(stratification, subclass = TRUE)
plot(balance.stratification2, var.order = "unmatched", abs = FALSE)
```

* Note that for some variables, while the groups are balanced in aggregate (black dots), the individual subclasses (gray numbers) may not be balanced, in which case unadjusted effect estimates within these subclasses should not be interpreted as unbiased.


### Calculate the stratum weights

To use bootstrap methods with the survey package, the surveyDesign object created above should be modified to include weights for each replication. The following code takes the surveyDesign object and adds weights for 1,000 bootstrapped samples:

```{r}
set.seed(8)
surveyDesignBoot <- as.svrepdesign(surveyDesign, type=c("bootstrap"), replicates=1000) 
```

The following R code uses svyby to apply the svymean function:

```{r}
head(lalonde)

subclassMeans <- svyby(formula=~re78, by=~treat+subclass,
                       design=surveyDesignBoot, FUN=svymean, covmat=TRUE)

subclassMeans
```

To obtain the ATE or ATT by pooling stratum-specific effects, svycontrast is used with the weights

First, use ntable to obtain the weights:

* For estimating the ATE, the stratum weight is wk = nk/n, which is the stratum size divided by the total sample size. 
* For estimating the ATT, the stratum weight is, wk = n1k/n1 which is the treated sample size within the stratum divided by the total treated sample size.

```{r}
subclass_table = balance.stratification2$qn[1:2,1:5]
ATEw = colSums(subclass_table)/sum(subclass_table)
ATTw = subclass_table[2,]/sum(subclass_table[2,])
subclass_table; ATEw; ATTw
```

This won't work: 

```{r, eval=FALSE}
pooledEffects <- svycontrast(subclassMeans, contrasts = list(ATT=ATTw)) 
```

* ATEw and ATTw needs to be specified for EVERY group within EVERY stratum
* in our case, 10 groups (2*5 = 10)
* Control groups get negative weights and treatment group gets positive weights: 

```{r}
subclassMeans$ATEsw <- c(-ATEw[1], ATEw[1], # first stratum, 
                         -ATEw[2], ATEw[2],  # second stratum
                         -ATEw[3], ATEw[3],  # third stratum
                         -ATEw[4], ATEw[4],  # fourth stratum
                         -ATEw[4], ATEw[5])  # fifth stratum
subclassMeans$ATTsw <- c(-ATTw[1], ATTw[1], # first stratum, 
                         -ATTw[2], ATTw[2],  # second stratum
                         -ATTw[3], ATTw[3],  # third stratum
                         -ATTw[4], ATTw[4],  # fourth stratum
                         -ATTw[4], ATTw[5])  # fifth stratum
subclassMeans
```

### Estimation of Treatment Effects

view ATTw2 here:

```{r}
pooledEffects <- svycontrast(subclassMeans, 
                             list(ATE = subclassMeans$ATTsw, ATT = subclassMeans$ATEsw)) 
pooledEffects
```

<!------------------------------>
## Adjustment
<!------------------------------>

```{r}
adjustModel <- lm(re78~treat+pscore,lalonde)
summ(adjustModel) 
```

<!------------------------------>
## Exercise
<!------------------------------>

Exercise: can you use fm2 to obtain propensity scores and see if that improves the balance and thus the estimate of treatment effect? 


