# Interupted Time Series (ITS) (2)

* Adapted from R codes in "Interrupted time series analysis using Autoregressive Integrated Moving Average (ARIMA) models: A guide for evaluating large-scale health interventions"
* Author: Dr. Andrea Schaffer

Install and load the following packages:

```{r, message=FALSE}
library(astsa)
library(forecast)
library(dplyr)
library(zoo)
library(tseries)
```

<!------------------------------>
## Data
<!------------------------------>

Load Data from csv:

```{r}
quet <- read.csv(file = 'quet.csv')
```

Convert data to time series object using ts():

```{r}
quet.ts <- ts(quet[,2], frequency=12, start=c(2011,1))
```

View data:

```{r}
quet.ts
```

<!------------------------------>
## Step 1: Plot data to visualize time series
<!------------------------------>

```{r}
options(scipen=5)
plot(quet.ts, xlim=c(2011,2015), ylim=c(0,40000), type='l', col="blue", 
     xlab="Month", ylab="Dispensings")
# Add vertical line indicating date of intervention (January 1, 2014)
abline(v=2014, col="gray", lty="dashed", lwd=2)
```


<!------------------------------>
## Step 2: Plot the differenced data
<!------------------------------>

seasonal differencing:

```{r}
plot(diff(quet.ts,12), xlim=c(2011,2015), ylim=c(-5000,5000), type='l',
     col="blue", xlab="Month", ylab="Dispensings")
# Add vertical line indicating date of intervention (January 1, 2014)
abline(v=2014, col="gray", lty="dashed", lwd=2)
```


first order differencing:

```{r}
plot(diff(diff(quet.ts,12)), xlim=c(2011,2015), ylim=c(-5000,5000), type='l',
     col="blue", xlab="Month", ylab="Dispensings")
# Add vertical line indicating date of intervention (January 1, 2014)
abline(v=2014, col="gray", lty="dashed", lwd=2)
```

second order differencing

```{r}
plot(diff(diff(diff(quet.ts,12))), xlim=c(2011,2015), ylim=c(-5000,5000), type='l', 
     col="blue", xlab="Month", ylab="Dispensings")
# Add vertical line indicating date of intervention (January 1, 2014)
abline(v=2014, col="gray", lty="dashed", lwd=2)
```

Test for stationary:

```{r}
# first order differencing
adf.test(diff(diff(quet.ts,12))) # not significant - not passed - nonstationary
# second order differencing
adf.test(diff(diff(diff(quet.ts,12)))) # significant - passed - stationary
```


<!------------------------------>
## Step 3: View ACF/PACF plots of differenced/seasonally differenced data
<!------------------------------>

```{r}
acf2(diff(diff(diff(quet.ts,12))), max.lag=24)
```

<!------------------------------>
## Step 4: Build ARIMA model
<!------------------------------>

1. Create variable representing step change and view

```{r}
step <- as.numeric(as.yearmon(time(quet.ts))>='Jan 2014')
step
```

2. Create variable representing ramp (change in slope) and view

```{r}
ramp <- append(rep(0,36), seq(1,12,1))
ramp  
```

* Use automated algorithm to identify p/q parameters
* Specify first difference = 2 and seasonal difference = 1

```{r, warning=FALSE, message=FALSE}
model1 <- auto.arima(quet.ts,                                 # data 
                     seasonal=TRUE,          # seasonal differencing 
                     xreg=cbind(step,ramp),    # intervention effect
                     d=2,                # non-seasonal differencing 
                     D=1,                    # seasonal differencing 
                     stepwise=FALSE,             # exhaustive search
                     trace=TRUE)                       # show output 
```


<!------------------------------>
## Step 5: Check residuals
<!------------------------------>

```{r}
checkresiduals(model1)
```

Estimate parameters and confidence intervals

```{r}
summary(model1)
confint(model1)
```

Plot data to visualize time series

```{r}
options(scipen=5)
plot(quet.ts, xlim=c(2011,2015), ylim=c(0,40000), type='l', col="blue", xlab="Month", ylab="Dispensings")

# Add vertical line indicating date of intervention (January 1, 2014)
abline(v=2014, col="gray", lty="dashed", lwd=2)
lines(fitted(model1), col="red", type="b")
```


<!------------------------------>
## Step 6: Calculate forecasts
<!------------------------------>

To forecast the counterfactual, model data excluding post-intervention time period

```{r}
model2 <- Arima(window(quet.ts, end=c(2013,12)), 
                order=c(2,2,1), 
                seasonal=list(order=c(0,1,1), period=12))
```

Forecast 12 months post-intervention and convert to time series object

```{r}
fc <- forecast(model2, h=12)
fc.ts <- ts(as.numeric(fc$mean), start=c(2014,1), frequency=12)
```

Combine with observed data

```{r}
quet.ts.2 <- ts.union(quet.ts, fc.ts)
```

Plot forecast

```{r}
plot(quet.ts.2, type="l", plot.type="s", col=c('blue','red'), xlab="Month", ylab="Dispensings", xlim=c(2011,2015), ylim=c(0,40000))
abline(v=2014, lty="dashed", col="gray")
```


