# Regression Discontinuity Designs (RDD)

* This lab goes over several estimators for RDD designs. 
* Part of the R codes were adapted from Chapter 13 of: R Companion to Real Econometrics: https://bookdown.org/carillitony/bailey/chp11.html#fn20

Install and load the following packages:

```{r, message=FALSE}
library(ggplot2)
library(tidyverse)
library(rdd)
```

<!------------------------------>
## Data
<!------------------------------>

* Let's estimate an RDD model using the data from a 2009 paper by Carpenter and Dobkin about the effect of increasing the drinking age on mortality rates. 
* Carpenter, Christopher and Carlos Dobkin. “The Effects of Alcohol Consumption on Mortality: Regression Discontinuity from the Minimum Drinking Age,” American Economic Journal: Applied Econometrics, 2009, 1:1, 164-182.
* The data are available at https://github.com/jrnold/masteringmetrics/tree/master/masteringmetrics/data in mlda.rda

* First, Let's look at motor vehicle deaths (outcome) as a function of age (assignment variable). The cutoff age is 21 years old. 


```{r}
load("mlda.rda")
mlda = na.omit(mlda) # removed 2 rows with missing data
head(mlda)
dim(mlda)
```

* assignment: minimum legal drinking age (agecell)
* cutoff: 21 years old 
* outcome: Motor Vehicle Accidents Death (mva) per 100,000 individuals 

Plot the outcome against assignment variable (ggplot2): 

```{r}
mlda %>% 
  select(agecell, mva) %>% 
  mutate(Treat = as.factor(ifelse(agecell >= 21, 1, 0))) %>% 
  ggplot(aes(x = agecell, y = mva, color = Treat)) +
  geom_point() + 
  geom_vline(xintercept = 21) + 
  labs(y = "Deaths in Moving Vehicle Accidents", x = "Age")
```

Alternatively, there is a generic way of plotting using plot(): 

```{r}
plot(mlda$agecell, mlda$mva, xlab = "Age", ylab = "Deaths in Moving Vehicle Accidents")
abline(v=21, col = 'red')
```

<!------------------------------>
## Segmented regression
<!------------------------------>

* There appears to be a discontinuity at age 21. Let’s estimate the RD model:
* mva = b0 + b1\*(agecell-21) + b2\*Treat + b3\*interaction1 + e
* where Treat=1 if agecell>=21 and Treat=0 if agecell<21
* let's call age_centered = agecell-21
* interaction1 = \*age_centered\*Treat

```{r}
mlda$Treat = ifelse(mlda$agecell>=21, 1, 0)  # dummy variable for policy
mlda$age_centered = mlda$agecell - 21  # "centered" age
mlda$int1 = mlda$Treat*mlda$age_centered   # interaction with "centered" age
```

Segmented regression:

```{r}
lm1 = lm(mva ~ age_centered + Treat + int1, data = mlda)
summary(lm1)
# interpretation?
```

Now, note that age must be centered at the cutoff for ease of interpretation: 

```{r}
mlda$int2 = mlda$Treat*mlda$agecell        # interaction with age

lm.wrong = lm(mva ~ agecell + Treat + int2, data = mlda)
summary(lm.wrong)
# interpretation?
```

```{r}
stargazer::stargazer(lm1, lm.wrong, type = 'text')
```

PLOTTING: it is easier and fancier to do the plotting using ggplot2, but it's also feasible to plot using generic plot()

```{r}
# RDD - unequal slope
mlda %>% 
  select(agecell, mva) %>% 
  mutate(Treat = as.factor(ifelse(agecell >= 21, 1, 0))) %>% 
  ggplot(aes(x = agecell, y = mva, color = Treat)) +
  geom_point() + 
  geom_smooth(method = "lm") + 
  labs(y = "Deaths in Moving Vehicle Accidents", x = "Age")
```


<!------------------------------>
## IK estimator
<!------------------------------>

Imbens-Kalyanaraman optimal bandwidth calculation:

```{r}
library(rdd)
rd1 = RDestimate(formula = mva ~ agecell, data = mlda, cutpoint = 21)
summary(rd1)
```


* In an RDD framework, we always need to balance our choice of large and small bandwidth. Larger bandwidth make the comparison between the treatment group and control group “less similar”.
* We see that the coefficient called “LATE”, is our parameter of interest of the MLDA program on mortality.
* We observe that deaths increase by 4.977 at the cutoff point of 21. The estimate of the treatment is slightly larger than the RDD estimate in lm1 (4.5340). 

smoothed curve:

```{r}
plot(rd1)
```

